---
output: 
  bookdown::pdf_document2:
    toc: false
    # number_sections: true
    includes:
      in_header: "Report.preamble.tex"
params: 
    sidebar_options: NULL
    surveyData: NULL
    surveyDataValidator: NULL
    surveyDataSummary: NULL
    nvcAssignment: NULL
    habCor: NULL
    floristicTables: NULL
    speciesFreq: NULL
    avgEIVs: NULL
    diversityAnalysis: NULL
    mvaNationalRefResults: NULL
    mvaLocalRefRestrictedResults: NULL
    mvaLocalRefUnrestrictedResults: NULL
    reportAuthorName: NULL
bibliography: Report.refs.bib
link-citations: yes
editor_options: 
  markdown: 
    wrap: 72
title: \textbf{\textcolor{ceh.dark.blue}{RMAVIS}}
subtitle: \textbf{\textcolor{ceh.dark.blue}{Report}}
# date: \textbf{\textcolor{ceh.dark.blue}{`r Sys.Date()`}}
---

```{r load_packages, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(kableExtra)
```

```{r retrieve_data, message=FALSE, warning=FALSE, include=FALSE}

# source("../create_constants.R", local = knitr::knit_global())
# sys.source("./create_constants.R", envir = knitr::knit_global())

# Retrieve report parameters
surveyData <- params$surveyData
surveyDataValidator <- params$surveyDataValidator
surveyDataSummary <- params$surveyDataSummary
nvcAssignment <- params$nvcAssignment
habCor <- params$habCor
sidebar_options <- params$sidebar_options
floristicTables <- params$floristicTables
floristicTables_composed_all <- floristicTables$floristicTables_composed_all
floristicTables_composed_all_wide <- floristicTables$floristicTables_composed_all_wide
speciesFreq <- params$speciesFreq
avgEIVs <- params$avgEIVs
diversityAnalysis <- params$diversityAnalysis
mvaNationalRefResults <- params$mvaNationalRefResults
mvaLocalRefRestrictedResults <- params$mvaLocalRefRestrictedResults
mvaLocalRefUnrestrictedResults <- params$mvaLocalRefUnrestrictedResults

# Survey data
surveyData_long <- surveyData$surveyData_long

# Sidebar options
inputMethod <- sidebar_options$inputMethod
habitatRestriction <- sidebar_options$habitatRestriction
nTopResults <- sidebar_options$nTopResults

# Report options
reportOptions <- sidebar_options$reportOptions
reportAuthorName <- sidebar_options$reportAuthorName
reportProjectName <- sidebar_options$reportProjectName

# Data Structure
surveyDataStructure <- surveyDataSummary$surveyDataStructure
quadratsPerYear <- surveyDataStructure$quadratsPerYear
quadratsPerYearGroup <- surveyDataStructure$quadratsPerYearGroup

# NVC Assignment
nvcAssignmentPlot_Jaccard <- nvcAssignment$nvcAssignmentPlot_Jaccard
nvcAssignmentSite_Czekanowski <- nvcAssignment$nvcAssignmentSite_Czekanowski
nvcAssignmentGroup_Czekanowski <- nvcAssignment$nvcAssignmentGroup_Czekanowski
topNVCCommunities <- nvcAssignment$topNVCCommunities

# Habitat Correspondences
# habCor <- habCor$

# EIVs
weightedMeanHEValuesSite <- avgEIVs$weightedMeanHEValuesSite
unweightedMeanHEValuesSite <- avgEIVs$unweightedMeanHEValuesSite
weightedMeanHEValuesGroup <- avgEIVs$weightedMeanHEValuesGroup
unweightedMeanHEValuesGroup <- avgEIVs$unweightedMeanHEValuesGroup
weightedMeanHEValuesQuadrat <- avgEIVs$weightedMeanHEValuesQuadrat
unweightedMeanHEValuesQuadrat <- avgEIVs$unweightedMeanHEValuesQuadrat

# Diversity
diversitySummary <- diversityAnalysis$diversitySummary
diversityIndices <- diversityAnalysis$diversityIndices
speciesRichnessSite <- diversityAnalysis$speciesRichnessSite
speciesRichnessGroup <- diversityAnalysis$speciesRichnessGroup
speciesRichnessQuadrat <- diversityAnalysis$speciesRichnessQuadrat

# MVA Local Reference, Restricted
mvaLocalRefRestricted_quadrats_survey <- mvaLocalRefRestrictedResults$dca_results_sample_site
mvaLocalRefRestricted_hulls_pquads <- mvaLocalRefRestrictedResults$pquad_hulls
mvaLocalRefRestricted_arrow_plot_data <- mvaLocalRefRestrictedResults$arrow_plot_data
mvaLocalRefRestricted_CCA_arrowData <- mvaLocalRefRestrictedResults$CCA_arrowData

# MVA Local Reference, Unrestricted
mvaLocalRefUnrestricted_quadrats_survey <- mvaLocalRefUnrestrictedResults$dca_results_sample_site
mvaLocalRefUnrestricted_hulls_pquads <- mvaLocalRefUnrestrictedResults$pquad_hulls
mvaLocalRefUnrestricted_arrow_plot_data <- mvaLocalRefUnrestrictedResults$arrow_plot_data
mvaLocalRefUnrestricted_CCA_arrowData <- mvaLocalRefUnrestrictedResults$CCA_arrowData

# MVA National Reference
mvaNationalRef_quadrats_survey <- mvaNationalRefResults$dca_results_sample_site
mvaNationalRef_hulls <- mvaNationalRefResults$pquad_hulls
mvaNationalRef_CCA_arrowData <- mvaNationalRefResults$CCA_arrowData


# Report options. Load from create_constants.R!!!
reportOptions_options <- list(`NVC Assignment` = c("Site, Czekanowski" = "nvcAssignmentResultsSite_Czekanowski",
                                                   "Group, Czekanowski" = "nvcAssignmentResultsGroup_Czekanowski",
                                                   "Quadrat, Jaccard" = "nvcAssignmentResultsQuadrat_Jaccard"),
                              `Habitat Correspondence` = c("Site" = "habitatCorrespondenceSite"),
                              `Floristic Tables` = c("Site" = "composedFloristicTablesSite",
                                                     "Group" = "composedFloristicTablesGroup"),
                              `Species Frequency` = c("Species Frequency" = "speciesFrequencyTable"),
                              `EIVs (incl. Mean Hill-Ellenberg)` = c("Weighted, Site" = "weightedMeanHEValuesSite",
                                                                     "Unweighted, Site" = "unweightedMeanHEValuesSite",
                                                                     "Weighted, Group" = "weightedMeanHEValuesGroup",
                                                                     "Unweighted, Group" = "unweightedMeanHEValuesGroup",
                                                                     "Weighted, Quadrat" = "weightedMeanHEValuesQuadrat",
                                                                     "Unweighted, Quadrat" = "unweightedMeanHEValuesQuadrat"),
                              `Diversity` = c("Summary" = "diversitySummary",
                                              "Quadrat Indices" = "diversityIndices",
                                              "Richness, Site" = "speciesRichnessSite",
                                              "Richness, Group" = "speciesRichnessGroup",
                                              "Richness, Quadrat" = "speciesRichnessQuadrat"),
                              `MVA` = c("National" = "mvaNationalReference",
                                        "Local (restricted)" = "mvaLocalReferenceRestricted",
                                        "Local (unrestricted)" = "mvaLocalReferenceUnrestricted"),
                              `Survey Table` = c("Survey Table" = "surveyData")
    )
```

\vspace{-10truemm}

```{r RMAVISVersion, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

cat("**Version:** ", as.character("v1.1.3"), sep = "")

```

```{r reportAuthorName, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

cat("**Name:** ", as.character(reportAuthorName), sep = "")

```

```{r reportProjectName, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

cat("**Project:** ", as.character(reportProjectName), sep = "")

```

```{r reportProjectDate, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

cat("**Date & Time:** ", as.character(format(Sys.time(), "%Y-%m-%d %H:%M:%S")), sep = "")

```

<!-- \newpage -->

\tableofcontents

\newpage

# Introduction

This report contains user-selected outputs of a RMAVIS application session.

## Main Options

```{r inputMethod, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
cat("**Input Method:** ", as.character(inputMethod), sep = "")
```

```{r habitatRestriction, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
cat("**Habitat Restriction:** ", as.character(habitatRestriction), sep = "")
```

```{r nTopResults, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
cat("**Number of Top Results:** ", as.character(nTopResults), sep = "")
```

## Selected Outputs

The report outputs selected by the app user are shown in Table \@ref(tab:selectedOutputs).

```{r selectedOutputs, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
reportOptions_options_named_df <- stack(lapply(reportOptions_options, names)) |>
  tibble::as_tibble() |>
  dplyr::select("Results Group" = ind, "Results Value" = values)

reportOptions_selected <- lapply(reportOptions_options, function(x) x[which(x %in% reportOptions)])
reportOptions_selected <- reportOptions_selected[lengths(reportOptions_selected) > 0L]
reportOptions_selected_df <- stack(lapply(reportOptions_selected, names)) |>
  tibble::as_tibble() |>
  dplyr::select("Results Group" = ind, "Results Value" = values) |>
  dplyr::mutate("Selected" = "Yes")

selectedOutputs <- reportOptions_options_named_df |>
  dplyr::left_join(reportOptions_selected_df, by = c("Results Group", "Results Value")) |>
  dplyr::mutate(
    "Selected" = dplyr::case_when(
      Selected == "Yes" ~ "Yes",
      TRUE ~ as.character("")
    )
  )

kableExtra::kbl(x = selectedOutputs, format = "latex", booktabs = TRUE,
                longtable = TRUE, linesep = "",
                caption = "The RMAVIS results selected for inclusion in this report.") |>
  kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))


```

\newpage

# Survey Data Structure

As detailed in the survey data validation section of RMAVIS
NVC similarities for groups and years are not calculated when there are
less than five sample plots per year-group and per year.
Smaller sample sizes may also result in the failure of the composition of 
representative floristic tables.
See Table \@ref(tab:quadratsPerYear) and Table \@ref(tab:quadratsPerYearGroup)
for the sample sizes present in the survey data.

```{r quadratsPerYear, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

kableExtra::kbl(x = quadratsPerYear, format = "latex", booktabs = TRUE,
                longtable = TRUE, linesep = "",
                caption = "The number of quadrats by year present in the inputted survey data.") |>
  kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))

```

```{r quadratsPerYearGroup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

kableExtra::kbl(x = quadratsPerYearGroup, format = "latex", booktabs = TRUE,
                longtable = TRUE, linesep = "",
                caption = "The number of quadrats by year and group present in the inputted survey data.") |>
  kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))

```

```{r nvcAssignmentTitle, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

reportOptions_nvcAssignment <- reportOptions_options$`NVC Assignment`

if(any(reportOptions_nvcAssignment %in% reportOptions)){
  
  cat("\\newpage")
  cat("# NVC Assignment")
  
}

```

```{r nvcAssignmentResultsSiteCzekanowski, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("nvcAssignmentResultsSite_Czekanowski" %in% reportOptions){
  
  cat("## NVC Assignment Czekanowski, by Site")
  
   if(!is.null(nvcAssignmentSite_Czekanowski)){
     
    nvcAssignmentSite_Czekanowski_prepped <- nvcAssignmentSite_Czekanowski |>
      dplyr::mutate_if(is.numeric, round, digits = 2)

    kableExtra::kbl(x = nvcAssignmentSite_Czekanowski_prepped, format = "latex", booktabs = TRUE,
                    longtable = TRUE, linesep = "",
                    caption = "The top NVC communities fitted using the Czekanowski method, for the site by year."
                    ) |>
      kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
     
   }
  
}

```

```{r nvcAssignmentResultsGroupCzekanowski, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("nvcAssignmentResultsGroup_Czekanowski" %in% reportOptions){
  
  cat("## NVC Assignment Czekanowski, by Group")
  
  if(!is.null(nvcAssignmentGroup_Czekanowski)){
    
    nvcAssignmentGroup_Czekanowski_prepped <- nvcAssignmentGroup_Czekanowski |>
      dplyr::mutate_if(is.numeric, round, digits = 2)

    kableExtra::kbl(x = nvcAssignmentGroup_Czekanowski_prepped, format = "latex", booktabs = TRUE,
                    longtable = TRUE, linesep = "",
                    caption = "The top NVC communities fitted using the Czekanowski method, by group and year"
                    ) |>
      kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
    
    
  }
  
}

```

```{r nvcAssignmentResultsQuadrat, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("nvcAssignmentResultsQuadrat_Jaccard" %in% reportOptions){
  
  cat("## NVC Assignment Jaccard, by Quadrat")
  
  if(!is.null(nvcAssignmentPlot_Jaccard)){
    
    nvcAssignmentPlot_Jaccard_prepped <- nvcAssignmentPlot_Jaccard |>
    dplyr::mutate_if(is.numeric, round, digits = 2)

    kableExtra::kbl(x = nvcAssignmentPlot_Jaccard_prepped, format = "latex", booktabs = TRUE,
                    longtable = TRUE, linesep = "",
                    caption = "The top NVC communities fitted using the pseudo-quadrat method, by quadrat, group, and year") |>
      kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
    
  }
  
}

```

```{r habitatCorrespondenceTitle, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("habitatCorrespondenceSite" %in% reportOptions){
  
  cat("# Habitat Correspondences")
  
}

```

```{r habitatCorrespondenceSite, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("habitatCorrespondenceSite" %in% reportOptions){
  
  cat("## Habitat Correspondences, by Site")
  
  habCor_prepped <- habCor |>
    dplyr::mutate_all(~replace(., is.na(.), ""))
  
  kableExtra::kbl(x = habCor_prepped, format = "latex", booktabs = TRUE,
                  longtable = TRUE, linesep = "",
                  caption = "The habitats associated with the top-N fitting NVC communities from alternative habitat classifications.") |>
    kableExtra::kable_styling(font_size = 7) |>
    kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position")) |>
    kableExtra::landscape()
  
}

```

```{r floristicTablesTitle, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

reportOptions_floristicTables <- reportOptions_options$`Floristic Tables`

if(any(reportOptions_floristicTables %in% reportOptions)){
  
  cat("\\newpage")
  cat("# Floristic Tables")
  
}

```

```{r floristicTablesSite, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("composedFloristicTablesSite" %in% reportOptions){
  
  cat("## Composed Floristic Tables, by Site")
  
  for(id in unique(floristicTables_composed_all$ID)){

    floristicTables_selected <- floristicTables_composed_all |>
      dplyr::filter(ID == id) |>
      dplyr::select(-ID)
  
    kableExtra::kbl(x = floristicTables_selected, format = "latex", booktabs = TRUE,
                    longtable = TRUE, linesep = "",
                    caption = paste0("Composed Floristic Table (", as.character(id), ")")) |>
      kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position")) |>
      print()
    
  }
  
}

```

```{r floristicTablesGroup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("composedFloristicTablesGroup" %in% reportOptions){
  
  cat("## Composed Floristic Tables, by Group")
  
  for(id in unique(floristicTables_composed_all$ID)){

    floristicTables_selected <- floristicTables_composed_all |>
      dplyr::filter(ID == id) |>
      dplyr::select(-ID)
  
    kableExtra::kbl(x = floristicTables_selected, format = "latex", booktabs = TRUE,
                    longtable = TRUE, linesep = "",
                    caption = paste0("Composed Floristic Table (", as.character(id), ")")) |>
      kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position")) |>
      print()
    
  }
  
}

```

```{r speciesFrequency, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("speciesFrequencyTable" %in% reportOptions){
  
  cat("\\newpage")
  cat("# Species Frequency")

  speciesFreqTable <- speciesFreq |>
    dplyr::mutate_all(~replace(., is.na(.), ""))
  
  kableExtra::kbl(x = speciesFreqTable, format = "latex", booktabs = TRUE,
                  longtable = TRUE, linesep = "",
                  caption = "The frequency of occurence of each species across all quadrats by year.") |>
    kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
  
}

```

```{r eivTitle, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

reportOptions_eiv <- reportOptions_options$`EIVs (incl. Mean Hill-Ellenberg)`

if(any(reportOptions_eiv %in% reportOptions)){
  
  cat("\\newpage")
  cat("# Ecological Indicator Values")
  
}

```

```{r weightedMeanHEValuesSite, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("weightedMeanHEValuesSite" %in% reportOptions){
  
  cat("## Weighted Mean Hill-Ellenberg Values, by Site")
  
  kableExtra::kbl(x = weightedMeanHEValuesSite, format = "latex", booktabs = TRUE,
                  longtable = TRUE, linesep = "",
                  caption = "Weighted Mean Hill-Ellenberg Values, by Site") |>
    kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
  
}

```

```{r unweightedMeanHEValuesSite, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("unweightedMeanHEValuesSite" %in% reportOptions){
  
  cat("## Unweighted Mean Hill-Ellenberg Values, by Site")
  
  kableExtra::kbl(x = unweightedMeanHEValuesSite, format = "latex", booktabs = TRUE,
                  longtable = TRUE, linesep = "",
                  caption = "Unweighted Mean Hill-Ellenberg Values, by Site") |>
    kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
  
}

```

```{r weightedMeanHEValuesGroup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("weightedMeanHEValuesGroup" %in% reportOptions){
  
  cat("## Weighted Mean Hill-Ellenberg Values, by Group")
  
  kableExtra::kbl(x = weightedMeanHEValuesGroup, format = "latex", booktabs = TRUE,
                  longtable = TRUE, linesep = "",
                  caption = "Weighted Mean Hill-Ellenberg Values, by Group") |>
    kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
  
}

```

```{r unweightedMeanHEValuesGroup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("unweightedMeanHEValuesGroup" %in% reportOptions){
  
  cat("## Unweighted Mean Hill-Ellenberg Values, by Group")
  
  kableExtra::kbl(x = unweightedMeanHEValuesGroup, format = "latex", booktabs = TRUE,
                  longtable = TRUE, linesep = "",
                  caption = "Unweighted Mean Hill-Ellenberg Values, by Group") |>
    kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
  
}

```

```{r weightedMeanHEValuesQuadrat, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("weightedMeanHEValuesQuadrat" %in% reportOptions){
  
  cat("## Weighted Mean Hill-Ellenberg Values, by Quadrat")
  
  kableExtra::kbl(x = weightedMeanHEValuesQuadrat, format = "latex", booktabs = TRUE,
                  longtable = TRUE, linesep = "",
                  caption = "Weighted Mean Hill-Ellenberg Values, by Quadrat") |>
    kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
  
}

```

```{r unweightedMeanHEValuesQuadrat, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("unweightedMeanHEValuesQuadrat" %in% reportOptions){
  
  cat("## Unweighted Mean Hill-Ellenberg Values, by Quadrat")
  
  kableExtra::kbl(x = unweightedMeanHEValuesQuadrat, format = "latex", booktabs = TRUE,
                  longtable = TRUE, linesep = "",
                  caption = "Unweighted Mean Hill-Ellenberg Values, by Quadrat") |>
    kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
  
}

```

```{r diversityTitle, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

reportOptions_diversity <- reportOptions_options$Diversity

if(any(reportOptions_diversity %in% reportOptions)){
  
  cat("\\newpage")
  cat("# Diversity")
  
}

```

```{r diversitySummary, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("diversitySummary" %in% reportOptions){
  
  cat("## Diversity Summary")
  
  kableExtra::kbl(x = diversitySummary, format = "latex", booktabs = TRUE,
                  longtable = TRUE, linesep = "",
                  caption = "Diversity Summary") |>
    kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
  
}

```

```{r diversityIndices, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("diversityIndices" %in% reportOptions){
   
  cat("## Diversity Indices")
  
  # cat("\\begin{landscape}")
  
  diversityIndices_prepped <- diversityIndices |>
    dplyr::mutate_if(is.numeric, round, digits = 2)
  
  kableExtra::kbl(x = diversityIndices_prepped, format = "latex", booktabs = TRUE,
                  longtable = TRUE, linesep = "",
                  caption = "Diversity Indices") |>
    # Note that "scale_down" does not work when "longtable = TRUE"
    # kableExtra::kable_styling(font_size = 7) |>
    kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position")) |>
    kableExtra::landscape()
  
   # cat("\\end{landscape}")
  
}

```

```{r speciesRichnessSite, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("speciesRichnessSite" %in% reportOptions){
  
  cat("## Species Richness, by Site")
  
  kableExtra::kbl(x = speciesRichnessSite, format = "latex", booktabs = TRUE,
                  longtable = TRUE, linesep = "",
                  caption = "Species Richness, by Site") |>
    kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
  
}

```

```{r speciesRichnessGroup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("speciesRichnessGroup" %in% reportOptions){
  
  cat("## Species Richness, by Group")
  
  kableExtra::kbl(x = speciesRichnessGroup, format = "latex", booktabs = TRUE,
                  longtable = TRUE, linesep = "",
                  caption = "Species Richness, by Group") |>
    kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
  
}

```

```{r speciesRichnessQuadrat, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("speciesRichnessQuadrat" %in% reportOptions){
  
  cat("## Species Richness, by Quadrat")
  
  kableExtra::kbl(x = speciesRichnessQuadrat, format = "latex", booktabs = TRUE,
                  longtable = TRUE, linesep = "",
                  caption = "Species Richness, by Quadrat") |>
    kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
  
}

```

```{r mvaTitle, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

reportOptions_mva <- reportOptions_options$MVA

if(any(reportOptions_mva %in% reportOptions)){
  
  cat("\\newpage")
  cat("# Multivariate Analysis")
  
}

```

```{r mvaNationalRefTitle, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.align = 'center', fig.cap='Multivariate analaysis plot, national reference.'}

if("mvaNationalReference" %in% reportOptions){
  
  # cat("\\newpage")
  cat("## National Reference")
  
}

```

```{r mvaNationalRef, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.align = 'center', fig.cap='Multivariate analaysis plot, national reference.'}

if("mvaNationalReference" %in% reportOptions){
  
  dcaAxisSelection <- "dca1dca2"
  x_axis <- "DCA1"
  y_axis <- "DCA2"
  
  mvaNationalRefPlot <- ggplot2::ggplot() +
    ggplot2::geom_polygon(data = mvaNationalRef_hulls |> dplyr::filter(dcaAxes == dcaAxisSelection,
                                                                       NVC %in% topNVCCommunities), 
                          alpha = 0.2,
                          mapping = ggplot2::aes(x = .data[[x_axis]],
                                                 y = .data[[y_axis]],
                                                 fill = NVC
                                                 )) +
    ggplot2::geom_point(data = mvaNationalRef_quadrats_survey,
                        mapping = ggplot2::aes(Year = Year,
                                               Group = Group,
                                               Quadrat = Quadrat,
                                               x = DCA1,
                                               y = DCA2)) +
    ggplot2::geom_segment(data = mvaNationalRef_CCA_arrowData,
                          color = 'black',
                          arrow = grid::arrow(),
                          mapping = ggplot2::aes(x = 0,
                                                 y = 0,
                                                 xend = CCA1,
                                                 yend = CCA2,
                                                 label = `Hill-Ellenberg`)) +
    ggplot2::geom_text(data = mvaNationalRef_CCA_arrowData,
                       color = 'black',
                       # position = ggplot2::position_dodge(width = 0.9),
                       size = 5,
                       mapping = ggplot2::aes(x = CCA1 * 1.075,
                                              y = CCA2 * 1.075,
                                              label = `Hill-Ellenberg`)) +
    ggplot2::theme_minimal()

  mvaNationalRefPlot
  
}

```

```{r mvaLocalRefRestrictedTitle, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.align = 'center', fig.cap='Multivariate analaysis plot, local reference, restricted.'}

if("mvaLocalReferenceRestricted" %in% reportOptions){
  
  cat("\\newpage")
  cat("## Local Reference, Restricted")
  
}

```

```{r mvaLocalRefRestricted, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.align = 'center', fig.cap='Multivariate analaysis plot, local reference, restricted.'}

if("mvaLocalReferenceRestricted" %in% reportOptions){
  
  dcaAxisSelection <- "dca1dca2"
  x_axis <- "DCA1"
  y_axis <- "DCA2"
  
  mvaLocalRefRestrictedPlot <- ggplot2::ggplot() +
    ggplot2::geom_polygon(data = mvaLocalRefRestricted_hulls_pquads  |> dplyr::filter(dcaAxes == dcaAxisSelection), 
                          alpha = 0.2,
                          mapping = ggplot2::aes(x = .data[[x_axis]],
                                                 y = .data[[y_axis]],
                                                 fill = NVC.Comm
                                                 )) +
    ggplot2::geom_point(data = mvaLocalRefRestricted_quadrats_survey,
                        mapping = ggplot2::aes(Year = Year,
                                               Group = Group,
                                               Quadrat = Quadrat,
                                               x = DCA1,
                                               y = DCA2)) +
    ggplot2::geom_segment(data = mvaLocalRefRestricted_CCA_arrowData,
                          color = 'black',
                          arrow = grid::arrow(),
                          mapping = ggplot2::aes(x = 0,
                                                 y = 0,
                                                 xend = CCA1,
                                                 yend = CCA2,
                                                 label = `Hill-Ellenberg`)) +
    ggplot2::geom_text(data = mvaLocalRefRestricted_CCA_arrowData,
                       color = 'black',
                       # position = ggplot2::position_dodge(width = 0.9),
                       size = 5,
                       mapping = ggplot2::aes(x = CCA1 * 1.075,
                                              y = CCA2 * 1.075,
                                              label = `Hill-Ellenberg`)) +
    ggplot2::theme_minimal()

  mvaLocalRefRestrictedPlot
  
}

```

```{r mvaLocalRefUnrestrictedTitle, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.align = 'center', fig.cap='Multivariate analaysis plot, local reference, unrestricted.'}

if("mvaLocalReferenceUnrestricted" %in% reportOptions){
  
  cat("\\newpage")
  cat("## Local Reference, Unrestricted")
  
}

```

```{r mvaLocalRefUnrestricted, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.align = 'center', fig.cap='Multivariate analaysis plot, local reference, unrestricted.'}

if("mvaLocalReferenceUnrestricted" %in% reportOptions){
  
  dcaAxisSelection <- "dca1dca2"
  x_axis <- "DCA1"
  y_axis <- "DCA2"
  
  mvaLocalRefUnrestrictedPlot <- ggplot2::ggplot() +
    ggplot2::geom_polygon(data = mvaLocalRefUnrestricted_hulls_pquads |> dplyr::filter(dcaAxes == dcaAxisSelection), 
                          alpha = 0.2,
                          mapping = ggplot2::aes(x = .data[[x_axis]],
                                                 y = .data[[y_axis]], 
                                                 fill = NVC.Comm
                                                 )) +
    ggplot2::geom_point(data = mvaLocalRefUnrestricted_quadrats_survey,
                        mapping = ggplot2::aes(Year = Year,
                                               Group = Group,
                                               Quadrat = Quadrat,
                                               x = DCA1,
                                               y = DCA2)) +
    ggplot2::geom_segment(data = mvaLocalRefUnrestricted_CCA_arrowData,
                          color = 'black',
                          arrow = grid::arrow(),
                          mapping = ggplot2::aes(x = 0,
                                                 y = 0,
                                                 xend = CCA1,
                                                 yend = CCA2,
                                                 label = `Hill-Ellenberg`)) +
    ggplot2::geom_text(data = mvaLocalRefUnrestricted_CCA_arrowData,
                       color = 'black',
                       # position = ggplot2::position_dodge(width = 0.9),
                       size = 5,
                       mapping = ggplot2::aes(x = CCA1 * 1.075,
                                              y = CCA2 * 1.075,
                                              label = `Hill-Ellenberg`)) +
    ggplot2::theme_minimal()

  mvaLocalRefUnrestrictedPlot
  
}

```

\newpage
# Appendix

```{r surveyData, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

if("surveyData" %in% reportOptions){
  
  cat("\\newpage")
  cat("## Survey Data Table")

  kableExtra::kbl(x = surveyData_long, format = "latex", booktabs = TRUE,
                  longtable = TRUE, linesep = "",
                  caption = "Survey Data Table") |>
    kableExtra::kable_styling(latex_options = c("repeat_header", "HOLD_position"))
  
}

```

