---
title: "pseudoMAVIS - Documentation"
subtitle: "v0.9"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    number_sections: true
    mathjax: "https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.5/MathJax.js?config=TeX-AMS_CHTML.js"
    self_contained: false
bibliography: references.bib
link-citations: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

```{css, echo=FALSE}

:root {
  
  --ceh-background-grey: #EBF0EE;
  --ceh-main-green: #4DA43B;
  --ceh-main-blue #2F7ECC:
  --ceh-main-darkblue #306FAE:
  --ceh-mid-1 #3889A3:
  --ceh-mid-2 #3F9280;
  --ceh-mid-3 #459A62;
  --text-title-color: #575656;
  
}

* {font-family: "Montserrat", sans-serif !important;}

body .main-container {
  max-width: 2800px !important;
  width: 1100px !important;
}

body {
  max-width: 2800px !important;
}

.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
  z-index: 2;
  color: black;
  background-color: var(--ceh-background-grey);
  border-color: var(--ceh-background-grey);
}

```

# Introduction

The Modular Analysis of Vegetation Information System
(MAVIS), originally developed by [@smart2016mavis]

## Aims

MAVIS aims to...

1.  Assist in the assignment of a set of sample quadrats to
    the closest NVC communities.
2.  Assist in the identification of site properties.
3.  ... This element of MAVIS is intended as an exploratory
    tool, not ....

## Scope

MAVIS is designed to analyse vegetation survey data in the
form of grouped or ungrouped quadrats, for a single site,
with or without cover estimates.

# Instructions and Methodology

## Survey Data

### Requirements

-   Each species value must match a species present in the
    accepted species list.
-   Each quadrat value must be unique, even if in a separate
    group.
-   Each year value must be a four digit integer.

Three methods for entering data into MAVIS are provided;
manual, upload, and example.

The list of species available for entry is currently
composed of:

-   The vascular plants present in @stace2010new, retrieved
    from @bsbi2014stace.
-   The bryophytes present in @blockeel2021.
-   The lichens, charophytes, and Algae present in the NVC
    floristic tables.

### Manual Entry

### Upload

### Example Data

Three example datasets are bundled with MAVIS, which provide
a means to explore the applications functionality.

**Parsonage Down** CG. Calcareous grassland. Transect
Survey. Repeat quadrats in 1970, 1990, and 2016
[@ridding2020]

**Leith** W. Bunce woodland Survey. Storm Track. Repeat
quadrats 1971, 2001, and 2021 [@smart2014]

**Whitwell Common** M. Norfolk Mires. Repeat quadrats 1994
and 1997. [@smart2000]

**Newborough Warren**. SD. Repeat quadrats. 2014 to 2018. [@wallace2018]

## NVC Assignment

1.  An R implementation of the Modular Analysis of
    Vegetation Information System (MAVIS). This method uses
    the Czekanowski index [@bloom1981] to quantify the
    similarity between a floristic table composed from
    survey data (with a minimum of five samples recommended)
    and the NVC Floristic Tables.

2.  An R implementation of the method outlined in
    @tipping2013. This method calculates the Jaccard
    similarity between user-inputted sample quadrats and
    pseudo-quadrats. This approach is noted as improving the
    accuracy of fit relative to the Czekanowski index for
    sample quadrats with poor species richness
    [@tipping2013].

## Habitat Correspondences

Habitat correspondence data is drawn from three sources:

1.  The Joint Nature Conservation Committee (JNCC)
    Spreadsheet of Habitat Correspondences [@jncc2008].
2.  For UKHab is drawn from UKHab V1.1 [@butcher2020].
3.  For the National Plant Monitoring Scheme (NPMS) data is
    drawn from [@pescott2019]

## Floristic Tables

To aid in the NVC assignment process and as an exploratory
tool the **Floristic Tables** section of MAVIS allows the
user to compare a composed floristic table

## Analysis Metrics

... uses the results of the NVC Assignment module to
calculate the following metrics:

-   Cover-weighted Hill-Ellenberg ecological indicator
    values:
    -   Nitrogen Score (N) [@bsbi2017ellenbergN]
    -   Moisture Score (F) [@bsbi2017ellenbergF]
    -   Reaction Score (R) [@bsbi2017ellenbergR]
    -   Salinity Score (S) [@bsbi2017ellenbergS]
    -   Light Score (L) [@bsbi2017ellenbergL]
-   

Data for bryophytes was taken from BRYOATT [@hill2007] and
data for vascular plants was taken from.

## Diversity

Metrics assessed:

-   Total Species Richness
-   Simpson's Diversity Index [@simpson1949].
-   Beta diversity [@koleff2003]

## Ordination

The ordination section of MAVIS aims to allow the user to
explore...

### Detrended Correspondence Analysis

Detrended Correspondence Analysis (DCA) [@hill1980dca]

### Canonical Correspondence Analysis

[@terBraak1986]

## Download Options

# Acknowledgements

# Author Contribution

Contributions towards the development of the application, in
accordance with [CRediT (Contributor Roles
Taxonomy)](https://credit.niso.org/), are:

**Zeke Marshall:** **Simon Smart:** **Colin Harrower:**

# Reccomended Citation

```{r bibtexRef}



```

# Release Log
<span style="font-size:2em;font-weight:700;">v0.9</span>

**2023/12/23**

-   First release for user testing.
-   Modifications:
-   Fixes:
-   New Features:


# References
