---
title: "Prepare Data for Bundling with MAVIS App"
output: html_document
date: "2023-11-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_constants}

source("create_constants.R")

```

## NVC Floristic tables

Prepare the NVC floristic tables.

```{r load_raw_NVCForTab, include=FALSE}

nvc_floristic_tables_raw <- read.csv(file = "../data/raw_data/NVC-floristic-tables.csv")

```

```{r create_NVCForTab, include=FALSE}

nvc_floristic_tables_tidied <- nvc_floristic_tables_raw |>
  dplyr::filter(is.na(Special.variable.value)) |>
  dplyr::select("NVC.Code" = Community.or.sub.community.code, 
                "Species" = Species.name.or.special.variable,
                "Constancy" = Species.constancy.value) |>
  dplyr::mutate("Species" = stringr::str_trim(Species))

```

```{r save_NVCForTab, include=FALSE}

saveRDS(object = nvc_floristic_tables_tidied, file = "../data/bundled_data/nvc_floristic_tables.rds")

```

Retrieve the NVC community codes.

```{r create_NVCCodes, include=FALSE}

nvc_name_to_code <- read.csv(file = "../data/raw_data/NVC-floristic-tables.csv") |>
  dplyr::select("Community.or.sub.community.name", "Community.or.sub.community.code") |>
  dplyr::distinct()

nvc_community_codes <- nvc_name_to_code |>
  dplyr::pull("Community.or.sub.community.code")

```

```{r save_NVCCodes, include=FALSE}

saveRDS(object = nvc_community_codes, file = "../data/bundled_data/nvc_community_codes.rds")

```

## Prepare bundled examples

```{r prepare_exampleData, include=FALSE}

dominCoverMidPointPerc_df <- tibble::enframe(dominCoverMidPointPerc) |>
  dplyr::rename("domin" = name,
                "Cover" = value)

example_data_df_all <- readRDS(file = "../data/raw_data/assignNVC_test_dataset.rds") |>
  dplyr::arrange(ID) |>
  dplyr::select("Sample" = ID, "Species" = species, "Site" = site, domin) |>
  dplyr::mutate("domin" = as.character(domin)) |>
  dplyr::left_join(dominCoverMidPointPerc_df, by = "domin") |>
  dplyr::select(-domin) |>
  dplyr::filter(Site != "")

exampleDataOptions <- example_data_df_all$Site |> unique()

setNames(exampleDataOptions, exampleDataOptions)

exampleDataOptions <- c("None" = "none", exampleDataOptions)

example_data_df_samples_per_site <- example_data_df_all |>
  dplyr::group_by(Site) |>
  dplyr::summarise(samples_per_site = dplyr::n())

example_data_df_trimmed <- example_data_df_all |>
  # Create a site sample number
  dplyr::group_by(Site) |>
  dplyr::mutate("Site.Sample" = as.integer(factor(Sample))) |>
  dplyr::ungroup() |>
  # Select only 5 samples per site as example data
  dplyr::filter(Site.Sample <= 5) |>
  dplyr::select(-Site.Sample)

```

Save the example data option names and example data.

```{r save_exampleData, include=FALSE}

saveRDS(object = exampleDataOptions, file = "../data/bundled_data/exampleDataOptions.rds")

saveRDS(object = example_data_df_trimmed, file = "../data/bundled_data/example_data_df.rds")

```

## BSBI BRC Codes Checklist

The BSBI BRC Codes checklist provides a concordance between plant taxon names, their BRC code, and the BSBI taxonId.
This concordance allows the use of all other checklist data.

```{r bsbiBRC_raw_data, include=FALSE}

sppName_to_brcCode_raw <- readr::read_delim("../data/raw_data/bsbi_checklist_sppName_to_brcCode.csv", 
                                            delim = "\t", escape_double = FALSE, 
                                            trim_ws = TRUE,
                                            show_col_types = FALSE)

```

```{r bsbiBRC_fix_data, include=FALSE}

sppName_to_brcCode <- sppName_to_brcCode_raw |>
  dplyr::group_by(dataValue) |>
  dplyr::filter(preferredTaxon == origTaxon) |>
  dplyr::rename("BRC" = "dataValue")

# If the length is equal there are no duplicate codes
nrow(sppName_to_brcCode) == length(sppName_to_brcCode$BRC |> unique())

# Find the duplicates
sppName_to_brcCode_duplicates <- sppName_to_brcCode |>
  dplyr::group_by(BRC) |>
  dplyr::filter(dplyr::n() > 1) |>
  dplyr::ungroup() |>
  dplyr::arrange(BRC)

# Remove duplicates
sppName_to_brcCode_NOduplicates <- sppName_to_brcCode |>
  dplyr::group_by(BRC) |>
  dplyr::filter(dplyr::n() == 1) |>
  dplyr::ungroup() |>
  dplyr::arrange(BRC)


# Align the BSBI BRC codes with assignNVC BRC codes
sppName_to_brcCode_alignedCodes <- sppName_to_brcCode_NOduplicates |>
  dplyr::mutate("BRC" = stringr::str_remove(string = BRC, pattern = "^0")) |>
  dplyr::mutate("BRC" = stringr::str_remove(string = BRC, pattern = "(?<=\\d)0+(?=0)")) |>
  dplyr::select(key, BRC, taxonId, preferredQualifiedTaxonName, preferredTaxon, preferredTaxonQualifier, origTaxon) |>
  # Shouldn't be any duplicates, there aren't.
  dplyr::distinct()

sppName_to_brcCode_alignedCodes_duplicates <- sppName_to_brcCode_alignedCodes |>
  dplyr::group_by(BRC) |>
  dplyr::filter(dplyr::n() > 1) |>
  dplyr::ungroup() |>
  dplyr::arrange(BRC)


# There are duplicate BRC codes due to families not adhering to the fix rule
# remove
sppName_to_brcCode_alignedCodes <- sppName_to_brcCode_alignedCodes |>
  dplyr::filter(!stringr::str_detect(preferredTaxon, pattern = "^[A-Za-z]+$"))

length(sppName_to_brcCode_alignedCodes$BRC) == nrow(sppName_to_brcCode_alignedCodes)

# sppName_to_brcCode_alignedCodes without BRC codes - should be a 0-row tibble
sppName_to_brcCode_alignedCodes_NOBRC <- sppName_to_brcCode_alignedCodes |>
  dplyr::filter(is.na(BRC)) |>
  dplyr::select(key) |>
  dplyr::distinct()

saveRDS(object = sppName_to_brcCode_alignedCodes, file = "../data/bundled_data/sppName_to_brcCode.rds")

```



## Pseudo-quadrat data

```{r ps_raw_data, include=FALSE}

nvc_pquads_raw <- assignNVC::nvc_pquads

```

Filter for unique species-BRC code cominations.
Note the the starting-number taxon group associations:
  - 2: Algae
  - 5: Lichens
  - 7: Charophytes
  - 8: Bryophytes
  - 9: Plants

```{r ps_speciesBRC_uniq, include=FALSE}

nvc_pquads_speciesBRC <- nvc_pquads_raw |>
  dplyr::select(species, BRC) |>
  dplyr::distinct()

```

There are a total of 22 species in `assignNVC::nvc_pquads` without BRC codes.

```{r ps_speciesBRC_noCodes, include=FALSE}

nvc_pquads_speciesBRC_noCodes <- nvc_pquads_speciesBRC |>
  dplyr::filter(is.na(BRC))

nvc_pquads_speciesBRC_noCodes

```

Manually add codes to missing species.

```{r ps_fixNoCodes, include=FALSE}

```

There are 112 observations in `assignNVC::nvc_pquads` which appear more than once.

```{r ps_duplicates, include=FALSE}

nvc_pquads_raw_dupes <- nvc_pquads_raw  |>
  dplyr::group_by(species, BRC, Pid2, NVC, Pid3, counter) |>
  dplyr::filter(dplyr::n() > 1) |>
  dplyr::distinct() |>
  dplyr::ungroup()

nvc_pquads_raw_dupes

```

Remove the duplicate observations.

```{r ps_NOduplicates, include=FALSE}

nvc_pquads_fixed <- nvc_pquads_raw |>
  dplyr::group_by(species, BRC, Pid2, NVC, Pid3, counter) |>
  # Remove duplicates
  dplyr::filter(dplyr::n() == 1) |>
  # Add duplicates
  rbind(nvc_pquads_raw_dupes)|>
  dplyr::ungroup() |>
  as.data.frame()

```

```{r save_ps, include=FALSE}

saveRDS(object = nvc_pquads_fixed, file = "../data/bundled_data/nvc_pquads_tidied.rds")

```


## Align Pseudo-quadrat data with BSBI BRC Codes

```{r align_ps_BSBIBRC, include=FALSE}

colnames(nvc_pquads_fixed) %in% colnames(sppName_to_brcCode_alignedCodes)
colnames(sppName_to_brcCode_alignedCodes) %in% colnames(nvc_pquads_fixed)

unique(sppName_to_brcCode_alignedCodes$BRC) |> length()

sppName_to_brcCode_alignedCodes_uniq <- unique(sppName_to_brcCode_alignedCodes$BRC)

nvc_pquads_joinedBSBIBRC <- nvc_pquads_fixed |>
  dplyr::mutate("BRC" = as.character(BRC)) |>
  dplyr::left_join(sppName_to_brcCode_alignedCodes, 
                   by = "BRC") |>
  dplyr::distinct()

# Joining the sppName_to_brcCode_alignedCodes should NOT add rows
nrow(nvc_pquads_joinedBSBIBRC) == nrow(nvc_pquads_fixed)

# Filter for pquads without BSBI data 
nvc_pquads_joinedBSBIBRC_none <- nvc_pquads_joinedBSBIBRC |>
  dplyr::filter(is.na(taxonId))

nvc_pquads_joinedBSBIBRC_none_spp <- nvc_pquads_joinedBSBIBRC_none |>
  dplyr::select(species, BRC) |>
  dplyr::distinct()

nvc_pquads_final <- nvc_pquads_joinedBSBIBRC

```


## BSBI checklist data

```{r prepare_BSBIChecklistData, include=FALSE}

HE_F <- readr::read_delim("../data/raw_data/bsbi_checklist_HE_F.csv", 
                          delim = "\t", escape_double = FALSE, 
                          trim_ws = TRUE,
                          show_col_types = FALSE) |>
  dplyr::select(key, taxonId, dataValue) |>
  dplyr::mutate("dataType" = "Hill-Ellenberg F")

HE_L <- readr::read_delim("../data/raw_data/bsbi_checklist_HE_L.csv", 
                          delim = "\t", escape_double = FALSE, 
                          trim_ws = TRUE,
                          show_col_types = FALSE) |>
  dplyr::select(key, taxonId, dataValue) |>
  dplyr::mutate("dataType" = "Hill-Ellenberg L")

HE_N <- readr::read_delim("../data/raw_data/bsbi_checklist_HE_N.csv", 
                          delim = "\t", escape_double = FALSE, 
                          trim_ws = TRUE,
                          show_col_types = FALSE) |>
  dplyr::select(key, taxonId, dataValue) |>
  dplyr::mutate("dataType" = "Hill-Ellenberg N")

HE_R <- readr::read_delim("../data/raw_data/bsbi_checklist_HE_R.csv", 
                          delim = "\t", escape_double = FALSE, 
                          trim_ws = TRUE,
                          show_col_types = FALSE) |>
  dplyr::select(key, taxonId, dataValue) |>
  dplyr::mutate("dataType" = "Hill-Ellenberg R")

HE_S <- readr::read_delim("../data/raw_data/bsbi_checklist_HE_S.csv", 
                          delim = "\t", escape_double = FALSE, 
                          trim_ws = TRUE,
                          show_col_types = FALSE) |>
  dplyr::select(key, taxonId, dataValue) |>
  dplyr::mutate("dataType" = "Hill-Ellenberg S")

rarityNewAtlas_GB <- readr::read_delim("../data/raw_data/bsbi_checklist_rarityNewAtlas_GB.csv", 
                                       delim = "\t", escape_double = FALSE, 
                                       trim_ws = TRUE,
                                       show_col_types = FALSE) |>
  dplyr::select(key, taxonId, dataValue) |>
  dplyr::mutate("dataType" = "Rarity - GB")

rarityNewAtlas_Ireland <- readr::read_delim("../data/raw_data/bsbi_checklist_rarityNewAtlas_Ireland.csv", 
                                            delim = "\t", escape_double = FALSE, 
                                            trim_ws = TRUE,
                                            show_col_types = FALSE) |>
  dplyr::select(key, taxonId, dataValue) |>
  dplyr::mutate("dataType" = "Rarity - Ireland")

bsbiChecklistData <- rbind(
  HE_F,
  HE_L,
  HE_N,
  HE_R,
  HE_S,
  rarityNewAtlas_GB,
  rarityNewAtlas_Ireland
)

```

```{r save_BSBIChecklistData, include=FALSE}

saveRDS(object = bsbiChecklistData, file = "../data/bundled_data/bsbiChecklistData.rds")

```


## Habitat correspondence data

```{r load_raw_habCorData, include=FALSE}

raw_JNCC_habCor <- readxl::read_xls(path = "../data/raw_data/Habitat-correspondences-2008.xls",
                                    sheet = "master table - sheet protected",
                                    col_types = c("text", "text", "text",
                                                  "text", "text", "text",
                                                  "text", "text", "text",
                                                  "text"))

ukHab_habCor_raw <- readxl::read_xlsx(path = "../data/raw_data/UK Habitat Classification V1-1 including Correspondences_7 Sep 2020_ZMCleaned.xlsx",
                                  sheet = "NVC to UKHab",
                                  skip = 6,
                                  .name_repair = "minimal")

ukHab_namesCodes_raw <- readxl::read_xlsx(path = "../data/raw_data/UK Habitat Classification V1-1 including Correspondences_7 Sep 2020_ZMCleaned.xlsx",
                                          sheet = "Professional Edition Hierarchy",
                                          skip = 0,
                                          .name_repair = "minimal")

```

```{r create_habCorData, include=FALSE}

ukHab_namesCodes_level2 <- ukHab_namesCodes_raw |>
  dplyr::select(`Level 2 code`, `Level 2 Label`) |>
  dplyr::rename("Code" = "Level 2 code", "Label" = "Level 2 Label") |>
  dplyr::filter(!is.na(Code)) |>
  dplyr::mutate("Level" = "2")

ukHab_namesCodes_level3 <- ukHab_namesCodes_raw |>
  dplyr::select(`Level 3 code`, `Level 3 Label`) |>
  dplyr::rename("Code" = "Level 3 code", "Label" = "Level 3 Label") |>
  dplyr::filter(!is.na(Code)) |>
  dplyr::mutate("Level" = "3")

ukHab_namesCodes_level4 <- ukHab_namesCodes_raw |>
  dplyr::select(`Level 4 code`, `Level 4 Label\r\n(Priority Habitats in Bold)`) |>
  dplyr::rename("Code" = "Level 4 code", "Label" = "Level 4 Label\r\n(Priority Habitats in Bold)") |>
  dplyr::filter(!is.na(Code)) |>
  dplyr::mutate("Level" = "4")

ukHab_namesCodes_level5 <- ukHab_namesCodes_raw |>
  dplyr::select(`Level 5 code`, `Level 5 Label\r\n(Including Annex 1 Habitats)`) |>
  dplyr::rename("Code" = "Level 5 code", "Label" = "Level 5 Label\r\n(Including Annex 1 Habitats)") |>
  dplyr::filter(!is.na(Code)) |>
  dplyr::mutate("Level" = "5")

ukHab_namesCodes <- rbind(
  ukHab_namesCodes_level2,
  ukHab_namesCodes_level3,
  ukHab_namesCodes_level4,
  ukHab_namesCodes_level5
)

ukHab_habCor_renamed <- ukHab_habCor_raw

names(ukHab_habCor_renamed) <- c("NVC.Code",
                                 "UKHab.Secondary.Code",
                                 "Level1_Variant1", "Level2_Variant1", "Level3_Variant1", "Level4_Variant1", "Level5_Variant1",
                                 "Level1_Variant2", "Level2_Variant2", "Level3_Variant2", "Level4_Variant2", "Level5_Variant2",
                                 "Level1_Variant3", "Level2_Variant3", "Level3_Variant3", "Level4_Variant3", "Level5_Variant3",
                                 "Level1_Variant4", "Level2_Variant4", "Level3_Variant4", "Level4_Variant4", "Level5_Variant4",
                                 "Level1_Variant5", "Level2_Variant5", "Level3_Variant5", "Level4_Variant5", "Level5_Variant5",
                                 "Level1_Variant6", "Level2_Variant6", "Level3_Variant6", "Level4_Variant6", "Level5_Variant6",
                                 "Level1_Variant7", "Level2_Variant7", "Level3_Variant7", "Level4_Variant7", "Level5_Variant7",
                                 "Level1_Variant8", "Level2_Variant8", "Level3_Variant8", "Level4_Variant8", "Level5_Variant8",
                                 "Level1_Variant9", "Level2_Variant9", "Level3_Variant9", "Level4_Variant9", "Level5_Variant9"
                                 )

ukHab_habCor_final <- ukHab_habCor_renamed |>
  dplyr::select(-UKHab.Secondary.Code) |>
  dplyr::mutate(dplyr::across(dplyr::everything(), as.character)) |>
  tidyr::pivot_longer(cols = !NVC.Code,
                      names_pattern = "(.*)_(.*)",
                      names_to = c("Level", "Variant"),
                      values_to = "Code") |>
  dplyr::distinct() |>
  tidyr::pivot_wider(names_from = Level,
                     values_from = Code) |>
  dplyr::filter(!is.na(Level1)) |>
  dplyr::mutate(
    Level2.Code = stringr::str_c(Level2),
    Level3.Code = stringr::str_c(Level2, Level3),
    Level4.Code = stringr::str_c(Level2, Level3, Level4),
    Level5.Code = stringr::str_c(Level2, Level3, Level4, Level5),
    
  ) |>
  dplyr::select(-c(Variant, Level1, Level2, Level3, Level4, Level5)) |>
  tidyr::pivot_longer(cols = c(Level2.Code, Level3.Code, Level4.Code, Level5.Code),
                      names_to = "Level",
                      values_to = "Code") |>
  dplyr::filter(!is.na(Code)) |>
  dplyr::mutate(
    Level = stringr::str_remove(string = Level,
                                pattern = stringr::fixed(".Code"))
  ) |>
  dplyr::distinct() |>
  dplyr::mutate("Relationship" = "Associated with", .after = "NVC.Code") |>
  dplyr::mutate("Classification" = paste0("UKHab - ", Level), .keep = "unused") |>
  dplyr::left_join(ukHab_namesCodes, by = "Code") |>
  dplyr::select(-Level) |>
  dplyr::select(NVC.Code, Relationship, Code, Label, Classification)


nvc_community_codes <- nvc_name_to_code |>
  dplyr::pull("Community.or.sub.community.code")

saveRDS(object = nvc_community_codes, file = "../data/bundled_data/nvc_community_codes.rds")

tidied_JNCC_habCor_CLASSN1 <- raw_JNCC_habCor |>
  dplyr::filter(CLASSN1 == "National Vegetation Classification") |>
  dplyr::select(-BIOTOPE_SHORT_TERM1,
                -CLASSN1,
                -`...10`) |>
  dplyr::relocate(RELATIONSHIP, .after = BIOTOPE_FULL_TERM1) |>
  dplyr::distinct()
  
tidied_JNCC_habCor_CLASSN1$BIOTOPE_FULL_TERM1 <- gsub("<[^>]+>", "", tidied_JNCC_habCor_CLASSN1$BIOTOPE_FULL_TERM1)
tidied_JNCC_habCor_CLASSN1$BIOTOPE_FULL_TERM2 <- gsub("<[^>]+>", "", tidied_JNCC_habCor_CLASSN1$BIOTOPE_FULL_TERM2)

jncc_habCor_final <- tidied_JNCC_habCor_CLASSN1 |>
  dplyr::select(BIOTOPE_CODE1, RELATIONSHIP, BIOTOPE_CODE2, BIOTOPE_FULL_TERM2, CLASSN2) |>
  dplyr::rename("NVC.Code" = "BIOTOPE_CODE1",
                "Relationship" = "RELATIONSHIP",
                "Code" = "BIOTOPE_CODE2",
                "Label" = "BIOTOPE_FULL_TERM2",
                "Classification" = "CLASSN2")

all_habCor_final <- rbind(
  ukHab_habCor_final,
  jncc_habCor_final
)

```

```{r save_habCorData, include=FALSE}

saveRDS(object = all_habCor_final, file = "../data/bundled_data/all_habCor_final.rds")

```

```{r create_habCorNames, include=FALSE}

all_habCor_classifications <- all_habCor_final |>
  dplyr::pull(Classification) |>
  unique()

```

```{r save_habCorNames, include=FALSE}

saveRDS(object = all_habCor_classifications, file = "../data/bundled_data/all_habCor_classifications.rds")

```


## Species names

Finally, create a list of accepted species names.

```{r create_speciesNames, include=FALSE}

speciesNames <- assignNVC::nvc_pquads |> # sppName_to_brcCode |>
  dplyr::pull(species) |>
  unique() |>
  sort()

```


```{r save_speciesNames, include=FALSE}

saveRDS(object = speciesNames, file = "../data/bundled_data/speciesNames.rds")

```

